Ages of Planets and Stars Observed by Kepler

### 2024-04-19 paper/ms.tex
update ms

\begin{figure*}[!t]
	\begin{center}
		\subfloat{
			\includegraphics[width=0.5\textwidth]{st_params_M_G_vs_dr3_bp_rp.pdf}
		}
		\subfloat{
			\includegraphics[width=0.5\textwidth]{st_params_M_G_vs_adopted_Teff.pdf}
		}

		\subfloat{
		\includegraphics[width=0.5\textwidth]{st_params_adopted_logg_vs_adopted_Teff.pdf}
		}
		\subfloat{
			\includegraphics[width=0.5\textwidth]{st_params_dr3_phot_g_mean_mag_vs_adopted_Teff.pdf}
		}
	\end{center}
	\vspace{-0.5cm}
	\caption{
    {\bf todo}.
	}
	\label{fig:stellarprops}
\end{figure*}

@@

@@

\subsection{Planet Quality Flags}
For the subset of our analysis that focuses on the properties of
planets, we use the following additional quality flags.

@@

\begin{figure*}[!t]
	\begin{center}
		\subfloat{
			\includegraphics[width=0.5\textwidth]{prot_teff_Santos19_Santos21_dquality.pdf}
		}
		\subfloat{
			\includegraphics[width=0.5\textwidth]{koi_mean_prot_teff_koi_X_S19S21dquality_keepgrazing.pdf}
		}
	\end{center}
	\vspace{-0.5cm}
	\caption{
    {\bf Rotation periods for Kepler stars (left) and planets (right)}.
    The periods are from \citet{Santos_2019} and \citet{Santos_2021}
    (\citetalias{Santos_2019} and \citetalias{Santos_2021}), and the
    temperatures are primarily from \citet{Berger_2020a_catalog}.  The
    gray lines are ``mean fits'' from \citet{Bouma_2023} to the slow
    rotation sequences of open clusters.  The stellar sample shows
    only apparently single stars near the main sequence with $\log
    g$$>$4.2, RUWE$<$1.4, 3800$<$$T_{\rm eff}$/K$<$6200, that are not
    flagged as eclipsing binaries.  The planet sample has the same
    stellar cuts, and shows confirmed and candidate planets from the
    cumulative KOI table, requiring them to have a MES$>$10.
	}
	\label{fig:prot_vs_teff}
\end{figure*}

@@

\begin{figure}[!t]
	\begin{center}
		\leavevmode
		\includegraphics[width=0.49\textwidth]{li_vs_teff_koi_X_S19S21dquality_eagles_showdispersion.pdf}
	\end{center}
	\vspace{-0.6cm}
	\caption{
    {\bf Equivalent widths (EW) of the \ion{Li}{1} 6708\,\AA\ doublet
    from Keck/HIRES spectra collected between 2009 and 2024.}
    Lines represent mean models derived from Gaia-ESO spectra of stars
    in clusters of various ages; the intrinsic dispersion 
    evolves in time, and is modelled as the $\pm$1$\sigma$
    bands~\citep{Jeffries_2023}.
    As for rotation-based ages, the intrisic dispersion sets the
    empirical precision limit for the technique.
		\label{fig:li_vs_teff}
	}
\end{figure}

>>>>

\begin{figure*}[!t]
	\begin{center}
		\subfloat{
			\includegraphics[width=0.47\textwidth]{st_params_M_G_vs_dr3_bp_rp.pdf}
		}
%	
%		\subfloat{
%			\includegraphics[width=0.4\textwidth]{st_params_M_G_vs_adopted_Teff.pdf}
%		}
%		\vspace{-0.5cm}
		\subfloat{
		\includegraphics[width=0.47\textwidth]{st_params_adopted_logg_vs_adopted_Teff.pdf}
		}
%		\subfloat{
%			\includegraphics[width=0.5\textwidth]{st_params_dr3_phot_g_mean_mag_vs_adopted_Teff.pdf}
%		}
	\end{center}
	\vspace{-0.5cm}
	\caption{
    {\bf The stellar samples.}  Our analysis focuses on stars observed by Kepler with 
    previously reported
    rotation periods (dark gray points). 
    About half of these rotators are suitable for gyrochronology (black points),
    based on factors including their temperatures, surface gravities, (non-)binarity, and proximity to the main sequence.
    The full list of criteria is given in Section~\ref{subsec:flags}.
    A few percent of the remainder are either ``confirmed'' or ``candidate'' Kepler Objects of Interest that
    meet an additional set of quality criteria (blue points; Section~\ref{subsec:plflags}).
    These are the objects for which our rotation-based ages are most reliable.
    We also report lithium-based ages for a largely-overlapping setting of KOIs
    based on high-resolution spectroscopy.
    Surface gravities and effective temperatures are primarily from
    \citet{Berger_2020a_catalog}.
	}
	\label{fig:stellarprops}
\end{figure*}

@@

\begin{figure*}[!t]
	\begin{center}
		\leavevmode
		\subfloat{
			\includegraphics[width=0.49\textwidth]{gyromodeldispersion.pdf}
		}
		\subfloat{
			\includegraphics[width=0.49\textwidth]{li_vs_teff_koi_X_S19S21dquality_eagles_showpoints_nodata.pdf}
		}
	\end{center}
	\vspace{-0.6cm}
	\caption{
		{\bf The models.}
		Points represent $10^4$ draws from model probability distributions
		that have been fitted to rotation periods 
		\citep{Bouma_2023} and to lithium measurements \citep{Jeffries_2023}
		of stars in open clusters.
		Lines are the fitted ``mean models'' at various ages; these models are calibrated
		for stars younger than 4\,Gyr.
		The intrinsic dispersion in rotation and lithium at any given age, which what these models are designed to fit, sets
		the theoretical precision limit for both age-dating methods.
		Additional sources of uncertainty, including measurement uncertainty,
		impose additional limitations on achievable age precision.
		The displayed points assume an unphysical uniform distribution in
		temperature for visual clarity, and sizes of the points are the same in each panel.
		%    Gyrochrone age posteriors are broadest in regions 
		%    with the slowest stellar spin-down.
		\label{fig:models}
	}
\end{figure*}

@@

\subsection{Planet Quality Flags}
\label{subsec:plflags}
For the subset of our analysis that focuses on the properties of
planets, we use the following additional quality flags.

@@

\begin{figure}[!t]
	\begin{center}
		\subfloat{
			\includegraphics[width=0.48\textwidth]{prot_teff_Santos19_Santos21_dquality.pdf}
		}

		\vspace{-0.5cm}
		\subfloat{
			\includegraphics[width=0.48\textwidth]{koi_mean_prot_teff_koi_X_S19S21dquality_keepgrazing.pdf}
		}
	\end{center}
	\vspace{-0.5cm}
	\caption{
    {\bf Rotation periods for Kepler stars (top) and planets (bottom)}.
    The periods are from \citet{Santos_2019} and \citet{Santos_2021}
    (\citetalias{Santos_2019} and \citetalias{Santos_2021}).  The
    gray lines are ``mean fits'' from Figure~\ref{fig:models} to the slow
    rotation sequences of M67, Ruprecht~147, NGC~6811, Praesepe, and a trio of 120\,Myr open clusters.
    The stellar sample shows
    only apparently single stars near the main sequence with $\log
    g$$>$4.2, RUWE$<$1.4, 3800$<$$T_{\rm eff}$/K$<$6200, that are not
    flagged as eclipsing binaries.  The planet sample has the same
    stellar cuts, and shows confirmed and candidate planets from the
    cumulative KOI table, requiring them to have a MES$>$10.
	}
	\label{fig:prot_vs_teff}
\end{figure}

@@

\begin{figure}[!t]
	\begin{center}
		\leavevmode
		\includegraphics[width=0.49\textwidth]{li_vs_teff_koi_X_S19S21dquality_eagles_logy.pdf}
	\end{center}
	\vspace{-0.6cm}
	\caption{
    {\bf Equivalent widths (EW) of the \ion{Li}{1} 6708\,\AA\ doublet
    from Keck/HIRES spectra collected between 2009 and 2024.}
    Lines are the ``mean'' isochrones from the same models as in Figure~\ref{fig:models}.
    The logarithmic scale highlights the dynamica range of both detections
    and upper limits.
%    The intrinsic
%    dispersion in the cluster data evolves in time, and is shown as
%    the $\pm$1$\sigma$ bands~\citep{Jeffries_2023}.
%    As for rotation-based ages, the intrisic dispersion sets the
%    empirical precision limit for the technique.
		\label{fig:li_vs_teff}
	}
\end{figure}
